---
title: "Cleaning Up  & reshaping merged electoral data into 'Long' format"
author: "Rebecca Linigen"
date: "23 September 2018"
output: html_document
---

#### I start by removing some un-needed columns in the data before trying to reshape the file

```{r, error=TRUE}
elec_result_PC_all <- select(elec_result_PC_all, -startsWith("DivisionID"))
```

#### I realised I hadn't reloaded the tidyverse library since my computer battery died

```{r, echo = TRUE, results='hide', warning=FALSE, error=FALSE, message=FALSE}
library(tidyverse)
library(tidyselect)
library(tidyr)
library(dplyr)
```

```{r, echo=TRUE, results='hide', warning=FALSE, error=FALSE, message=FALSE}
elec_result_PC_all <- read_csv("/Users/Bec/R Projects/AT2/cleaned_data/elec_results_PC_all.csv")
```

```{r, error= TRUE}
glimpse(elec_result_PC_all)
```

```{r, error=TRUE}
elec_result_PC_all <- select_(elec_result_PC_all, -startsWith("DivisionID"))
```

#### ok let's try that again, I'm going to test this on 'DivisionID' - it looks like `starts_with` needed the underscore

```{r,error=TRUE} 
elec_result_PC_all <- select(elec_result_PC_all, -starts_with("DivisionID"))
```

#### Thats worked, ok now for more columns

```{r,error=TRUE} 
elec_result_PC_all <- select(elec_result_PC_all, -starts_with("DivisionID", "Surname", "GivenNm", "PartyAB", "CandidateID"))
```

#### ok it looks like I can't do multiple at once this way, I'm going to try `one-of` instead.

```{r, error=TRUE, warning=FALSE} 
elec_result_PC_all <- select(elec_result_PC_all, -one_of(c("DivisionID", "CandidateID")))
```

#### That worked, but seems a bit tedious naming every single column seperately, there must be a better way, back to `starts_with`, at least that can remove column 'groups' with the same start. 

```{r, error =TRUE} 
elec_result_PC_all <- select(elec_result_PC_all, -starts_with("DivisionID", "Surname", "GivenNm", "PartyAb", "CandidateId"))
```

#### No luck, how about `contains`

```{r, error=TRUE} 
elec_result_PC_all <- select(elec_result_PC_all, -contains("DivisionID", "Surname", "GivenNm", "PartyAb", "CandidateId"))
```

#### Back to `starts_with` but why does this only seem to work one by one (?), this will take a while

```{r,error=TRUE} 
elec_result_PC_all <- select(elec_result_PC_all, -starts_with("Surname"))
elec_result_PC_all <- select(elec_result_PC_all, -starts_with("GivenNm"))
elec_result_PC_all <- select(elec_result_PC_all, -starts_with("PartyAb"))
```

#### How about a few at once

```{r,error =TRUE} 
elec_result_PC_all <- select(elec_result_PC_all, -starts_with("Surname", "GivenNm"))
```
#### Didn't like that. The last column to clean up is state, but I want to leave one of them behind, the other 2 can go (we don't need 3 state columns by year seeing as they don't change each election)

```{r, error=TRUE, eval=FALSE} 
rename(elec_result_PC_all, "State" = "StateAb.x")
```

####This looks promising, let's take a look

```{r, error=TRUE} 
head(elec_result_PC_all)
```
#### Now to remove those extra state by year columns.

```{r, error=TRUE} 
elec_result_PC_all <- select(elec_result_PC_all, -matches("StateAb.y"))
elec_result_PC_all <- select(elec_result_PC_all, -matches("StateAb"))
```

#### And the final piece which is the CandidateID columns
```{r, error=TRUE} 
elec_result_PC_all <- select(elec_result_PC_all, -starts_with("CandidateID"))
```

#### Now that I have a cleaner data file in `wide` format, I'd like to convert it to `long` format. Before I do this, I ned to make sure the postcode column is a factor.
```{r, error=TRUE, eval = FALSE} 
elec_result_PC_all$Postcode <- factor(elec_result_PC_all$Postcode)
```
#### I found [this Cookbook for R article](http://www.cookbook-r.com/Manipulating_data/Converting_data_between_wide_and_long_format/) really helpful in showing me how to reshape data using the `gather` function. 

#### Now use the `gather` function to combine 'PartyNm2016', 'PartyNm2013', and PartyNm2010' into the column 'Party_Year'
```{r, error=TRUE, eval=FALSE}
elec_result_PC_all_long <- gather(elec_result_PC_all, Party_Year, Party_Name, PartyNm2016:PartyNm2013, factor_key = TRUE)
```

```{r, error=TRUE, eval=FALSE, warning =FALSE}
head(elec_result_PC_all_long)
```

#### Rename factor names from "PartyNm2016", "PartyNm2010" and "PartyNm2013" to "2016", "2010" and "2013"
```{r, error=TRUE, eval =FALSE} 
levels(elec_result_PC_all_long$Party_Year)[levels(elec_result_PC_all_long$Party_Year)=="PartyNm2016"] <- "2016"
levels(elec_result_PC_all_long$Party_Year)[levels(elec_result_PC_all_long$Party_Year)=="PartyNm2013"] <- "2013"
levels(elec_result_PC_all_long$Party_Year)[levels(elec_result_PC_all_long$Party_Year)=="PartyNm2010"] <- "2010"
```

## Done - ready for merging with the other datasets in `long` format & with unecessary columns removed. 
